Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-04 paper/Conclusion.tex
some modifications of backgrounda

\section{Discussion}

We illustrated a direct realization of our proposed proof framework
for a very flexible cache coherence protocol implementation codifying
the intuition behind the system's correctness as simple
invariants. While this was took significant expertise, once done, we
were able to show that many key hardware modifications,\eg\ caching
policies, could be made with zero or very small natural changes to the
proof. These bits were small enough that an informed hardware designer
implementing our protocols can reasonably be expected to provide such
information during refinement giving them a viable path for producing
a provably correct performant cache coherence engine implementation.

The protocol framework we showed was very flexible, and could easily
be extended for variants that go beyond microarchitectural variants,
\eg\ silent drops, extension to MOSI with a relatively small amount of
additional reasoning.  This gives us good confidence that we can
effectively produce frameworks for other classes protocol,
\eg\ update-based protocols, non-inclusive cache hierarchies. Further,
nothing about our methodology requires we limit ourselves to cache
coherence systems; this approach should naturally extend to any
hardware-based system.

While fundamentally this work meets our goal of carrying verification
into through the design refinement step, there are still some
important mechanization to make this ready for industrial
use. Foremost, the simple transliteration between the Coq transition
system and the Bluespec hardware description implementation should be
fully mechanized. Further, the proposed extension to provide fairness
of transaction scheduling in the Bluespec compiler should be
implemented. Finally, to fully guarantee correctness of the resulting
hardware FSM, we should also fully verify the hardware generation as
in the FeSi~\cite{Tomas} efforts.

>>>>

\section{Conclusions and future work}

We illustrated a direct realization of our proposed proof framework for a
concrete MSI cache coherence protocol implementation codifying the intuition
behind the system's correctness as simple invariants. While this took
significant effort, once done, we showed that many key hardware
modifications,\eg{} replacement policies, could be made with a very small
addition to the proof. These bits were small enough that an informed hardware
designer implementing our protocols can reasonably be expected to provide such
additional proofs during refinement, producing a provably correct realistic
cache coherence engine implementation.

The protocol framework we showed was very flexible, and could easily be
extended for variants that go beyond micro-architectural variants, \eg{}
extension to MOSI with a relatively small amount of additional reasoning. This
gives us good confidence that we can effectively produce frameworks for other
classes of protocols, \eg{} update-based protocols, non-inclusive cache
hierarchies. Further, nothing about our methodology requires we limit ourselves
to cache coherence systems; this approach should naturally extend to any
hardware-based system.

While fundamentally this work meets our goal of verification of synthesizable
hardware specifications, there is still some important mechanization to make
this ready for industrial use. Foremost, the simple transliteration between the
inductively defined transition system in Coq and the Bluespec hardware
description should be fully mechanized.
%Further, the proposed extension to provide fairness of transaction scheduling
%in the Bluespec compiler should be implemented.
Further, to fully guarantee correctness of the resulting hardware, we should
also verify the hardware synthesis compiler, possibly using FeSi~\cite{Tomas}.
